\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control for Energy and Thermal Comfort Optimization in Building Energy Systems using Random Forests}

### 2018-01-20 paper/final/cs-dpc.tex
small changes

The fundamental reason why DPC  is suitable for such a problem is that when the complexity rises, there is a huge cost to model all the states given by the dynamical system \eqref{E:bilinear1}. For example, states in the bilinear model also include slab temperatures which require modeling of structural and material properties in detail and often we also need to install new sensors to capture additional states. Thus, DPC is based solely on one state of the model, i.e. the zone temperature that can be easily measured with a thermostat. This serves as the output variable in $\Y$ of interest for which we build $N$ trees and $N$ forests as described in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively. Therefore, $y_{k+j}:=x_{k+j+1}^1$, where $x^1$ is the first component of $x$.
Next, we define the non-manipulated features in $\X^d$. At time $k$, for the tree $\mathcal{T}_j$ and the forest $\mathcal{F}_j$, we base these features to include weather disturbances, external disturbances due to occupancy and equipments, and autoregressive terms of the room temperature, i.e.
$\X^d:=\{( d_{k+j-N} ,\ldots,d_{k+j-1}, x_{k}^1,\ldots, x_{k-\delta}^1 )\},\ \forall k=1,\ldots,n$, where $\delta$ is the order of state autoregression.
Finally, the inputs in DPC are exactly same as in MPC, i.e. $\X^c:=u_k,\ \forall k=1,\ldots,n$.
The training data in the above format was generated by simulating the bilinear model with rule-based strategies for 10 months in 2007. January and May were deliberately excluded for testing the DPC implementation.
\subsubsection{Optimization} 
\label{SSS:dpc_opt}
For a fair comparison with MPC, we cast DPC optimization problem as follows:

\begin{subequations}
	\begin{align}
	& \underset{u_{k+j},\epsilon_j}{\text{minimize}} & & \sum_{j=1}^{N} (y_{k+j} - x_{ref}) \mathcal{Q}^{(1,1)} (y_{k+j} - x_{ref}) + c^T u_{k+j-1} + \lambda\epsilon_j \\
	& \text{subject to }                  & & y_{k+j} =  \alpha_j \left[1,u_{k},\ldots,u_{k+j-1} \right]^\top \label{SE:dpc}                                            \\
	&                                     & & \underline{u} \leq u_{k+j-1} \leq \bar{u} 									   	                                        \\
	&                                     & & \underline{y} - \epsilon_j \leq y_{k+j} \leq \bar{y} + \epsilon_j                                                         \\
	&                                     & & \epsilon_j \geq 0, \ j = 1,\dots,N,            									                 
	\end{align}\label{E:dpc}
\end{subequations}

@@

We compare the prediction for the first time step $y_{k+1}$ and the 6-hour ahead prediction $y_{k+6}$, given $k$, for a week in the month of May in Figure~\ref{F:validation}. It is visible how trees have a high variance, and the forests are more accurate. Note that data from January and May were not used for training. The quantitative summary of the accuracy is given in Table~\ref{T:validation}. We can see that the random forests are better in all respects.
\begin{table}[h!]
	\centering
	\begin{tabular}{cccc}
		\toprule
		                  & RMSE & $R^2$ score & EV  \\ 
		\midrule
		tree-$y_{k+1}$    & 0.42 & 0.75        & 0.76 \\
		tree-$y_{k+6}$    & 0.64 & 0.41        & 0.42 \\
		forest-$y_{k+1}$  & 0.29 & 0.87        & 0.88 \\
		forest-$y_{k+6}$  & 0.38 & 0.78        & 0.80 \\
		\bottomrule
	\end{tabular}
	\caption{Quantitative comparison of root mean square error (RMSE), $R^2$ score, and explained variance (EV) for trees and forests for different predictions steps.}
	\captionsetup{justification=centering}
	\label{T:validation}
\end{table}

>>>>

The fundamental reason why DPC  is suitable for such a problem is that when the complexity rises, there is a huge cost to model all the states given by the dynamical system \eqref{E:bilinear1}. For example, states in the bilinear model also include slab temperatures which require modeling of structural and material properties in detail and often we also need to install new sensors to capture additional states. Thus, DPC is based solely on one state of the model, i.e. the zone temperature that can be easily measured with a thermostat. This serves as the output variable in $\Y$ of interest for which we build $N$ trees and $N$ forests as described in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively. Therefore, $y(k+j):=x(k+j+1)^1$, where $x^1$ is the first component of $x$.
Next, we define the non-manipulated features in $\X^d$. At time $k$, for the tree $\mathcal{T}_j$ and the forest $\mathcal{F}_j$, we base these features to include weather disturbances, external disturbances due to occupancy and equipments, and autoregressive terms of the room temperature, i.e.
$\X^d:=\{( d(k+j-N) ,\ldots,d(k+j-1), x(k)^1,\ldots, x(k-\delta)^1 )\},\ \forall k=1,\ldots,n$, where $\delta$ is the order of state autoregression.
Finally, the inputs in DPC are exactly same as in MPC, i.e. $\X^c:=\{u(k)\},\ \forall k=1,\ldots,n$.
The training data in the above format was generated by simulating the bilinear model with rule-based strategies for 10 months in 2007. January and May were deliberately excluded for testing the DPC implementation.
\subsubsection{Optimization} 
\label{SSS:dpc_opt}
For a fair comparison with MPC, we cast DPC optimization problem as follows:

\begin{subequations}
	\begin{align}
	& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} (y_{k+j}-x_{ref})\mathcal{Q}^{(1,1)} (y_{k+j}-x_{ref})+c^\top u_{k+j}+\lambda\epsilon_j \\
	& \mathrm{subject\ to }                              & & y_{k+j} =  \alpha_j \left[1,u^\top_{k},\ldots,u^\top_{k+j} \right]^\top \label{SE:dpc}               \\
	&                                                    & & \underline{u} \leq u_{k+j} \leq \bar{u} 	                                                          \\
	&                                                    & & \underline{y} - \epsilon_j \leq y_{k+j} \leq \bar{y} + \epsilon_j                                    \\
	&                                                    & & \epsilon_j \geq 0, \ j = 0,\dots,N,            									                 
	\end{align}\label{E:dpc}
\end{subequations}

@@

We compare the prediction for the first time step $y(k+1)$ and the 6-hour ahead prediction $y(k+6)$, given $k$, for a week in the month of May in Figure~\ref{F:validation}. It is visible how trees have a high variance, and the forests are more accurate. Note that data from January and May were not used for training. The quantitative summary of the accuracy is given in Table~\ref{T:validation}. We can see that the random forests are better in all respects.
\begin{table}[h!]
	\centering
	\begin{tabular}{lccc}
		\toprule
		                 & RMSE & $R^2$ score & EV  \\ 
		\midrule
		tree - $y(k+1)$    & 0.42 & 0.75        & 0.76 \\
		tree - $y(k+6)$    & 0.64 & 0.41        & 0.42 \\
		forest - $y(k+1)$  & 0.29 & 0.87        & 0.88 \\
		forest - $y(k+6)$  & 0.38 & 0.78        & 0.80 \\
		\bottomrule
	\end{tabular}
	\caption{Quantitative comparison of root mean square error (RMSE), $R^2$ score, and explained variance (EV) for trees and forests for different predictions steps.}
	\captionsetup{justification=centering}
	\label{T:validation}
\end{table}

### 2018-01-20 paper/final/cs-model.tex
small changes

The bilinear model has 12 internal states including the inside zone temperature $\mathsf{T}_{\mathrm{in}}$, the slab temperatures $\mathsf{T}_{\mathrm{sb}}$, the inner wall $\mathsf{T}_{\mathrm{iw}}$ and the outside wall temperature $\mathsf{T}_{\mathrm{ow}}$. The state vector is defined as $x:=[\mathsf{T}_{\mathrm{in}}, \mathsf{T}_{\mathrm{sb}}^{(1:5)}, \mathsf{T}_{\mathrm{ef}}^{(1:3)}, \mathsf{T}_{\mathrm{in}}^{(1:3)}]^T$.

There are 4 control inputs including the blind position $\mathsf{B}$, the gains due to electric lighting $\mathsf{L}$, the evaporative cooling usage factor $\mathsf{C}$, and the heat from the radiator $\mathsf{H}$ such that $u:=[\mathsf{B},\mathsf{L},\mathsf{H},\mathsf{C}]^T$. $\mathsf{B}$ and $\mathsf{L}$ affect both room illuminance and temperature due to heat transfer whereas $\mathsf{C}$ and $\mathsf{H}$ affect only temperature.

The model is subject to 5 weather disturbances: solar gains with fully closed blinds $\mathsf{Q}_{\mathrm{sc}}$ and with open blinds $\mathsf{Q}_{\mathrm{so}}$, daylight illuminance with open blinds $\mathsf{I}_{\mathrm{o}}$, external dry-bulb temperature $\mathsf{T}_{\mathrm{db}}$ and external wet-bulb temperature $\mathsf{T}_{\mathrm{wb}}$. 
The hourly weather forecast, provided by MeteoSwiss, was updated every 12 hrs. Therefore, to improve the forecast,  an autoregressive model of the uncertainty was considered.
Other disturbances come from the internal gains due to occupancy $\mathsf{Q}_{\mathrm{io}}$ and due to equipments $\mathsf{Q}_{\mathrm{ie}}$ which were assumed as per the Swiss standards \cite{Merkblatt2006}. We define $d:=[\mathsf{Q}_{\mathrm{sc}},\mathsf{Q}_{\mathrm{so}},\mathsf{I}_{\mathrm{o}},\mathsf{Q}_{\mathrm{io}},\mathsf{Q}_{\mathrm{ie}},\mathsf{T}_{\mathrm{db}},\mathsf{T}_{\mathrm{wb}}]^T$. For further details, we refer the reader to \cite{Oldewurtel2011}.

The model dynamics are given below. The bilinearity is present in both input-state, and input-disturbance.
\begin{equation}\label{E:bilinear1}
x_{k+1} = Ax_{k}+(B_u +B_{xu}[x_k] + B_{du}[d_k]) u_k+B_dd_k,
\end{equation}
where $x_{k} \in \mathbb{R}^{12}, u_{k} \in \mathbb{R}^{4}, d_{k} \in \mathbb{R}^{8} \ \forall k = 0,\dots,T$, and the matrices $B_{xu}$ and $B_{du}$ are defined as
\begin{equation}
	\begin{aligned}
		&B_{xu}[x_k] = [ B_{xu,1}[x_k],B_{xu,2}[x_k], \dots, B_{xu,4}[x_k] ] \in \mathbb{R}^{12\times4}, \\
		&B_{du}[d_k] = [ B_{du,1}[d_k],B_{du,2}[d_k], \dots, B_{du,4}[d_k] ] \in \mathbb{R}^{12\times4}, \\
	\end{aligned}
\end{equation}
with $B_{xu,i} \in \R^{12\times12}, B_{du,i} \in \R^{12\times8} \ \forall i=1,2,3,4$.
For this study, we assume that the disturbances are precisely known to MPC as well as DPC controller.

>>>>

The bilinear model has 12 internal states including the inside zone temperature $\mathsf{T}_{\mathrm{in}}$, the slab temperatures $\mathsf{T}_{\mathrm{sb}}$, the inner wall $\mathsf{T}_{\mathrm{iw}}$ and the outside wall temperature $\mathsf{T}_{\mathrm{ow}}$. The state vector is defined as $x:=[\mathsf{T}_{\mathrm{in}}, \mathsf{T}_{\mathrm{sb}}^{(1:5)}, \mathsf{T}_{\mathrm{ef}}^{(1:3)}, \mathsf{T}_{\mathrm{in}}^{(1:3)}]^\top$.

There are 4 control inputs including the blind position $\mathsf{B}$, the gains due to electric lighting $\mathsf{L}$, the evaporative cooling usage factor $\mathsf{C}$, and the heat from the radiator $\mathsf{H}$ such that $u:=[\mathsf{B},\mathsf{L},\mathsf{H},\mathsf{C}]^\top$. $\mathsf{B}$ and $\mathsf{L}$ affect both room illuminance and temperature due to heat transfer whereas $\mathsf{C}$ and $\mathsf{H}$ affect only temperature.

The model is subject to 5 weather disturbances: solar gains with fully closed blinds $\mathsf{Q}_{\mathrm{sc}}$ and with open blinds $\mathsf{Q}_{\mathrm{so}}$, daylight illuminance with open blinds $\mathsf{I}_{\mathrm{o}}$, external dry-bulb temperature $\mathsf{T}_{\mathrm{db}}$ and external wet-bulb temperature $\mathsf{T}_{\mathrm{wb}}$. 
The hourly weather forecast, provided by MeteoSwiss, was updated every 12 hrs. Therefore, to improve the forecast,  an autoregressive model of the uncertainty was considered.
Other disturbances come from the internal gains due to occupancy $\mathsf{Q}_{\mathrm{io}}$ and due to equipments $\mathsf{Q}_{\mathrm{ie}}$ which were assumed as per the Swiss standards \cite{Merkblatt2006}. We define $d:=[\mathsf{Q}_{\mathrm{sc}},\mathsf{Q}_{\mathrm{so}},\mathsf{I}_{\mathrm{o}},\mathsf{Q}_{\mathrm{io}},\mathsf{Q}_{\mathrm{ie}},\mathsf{T}_{\mathrm{db}},\mathsf{T}_{\mathrm{wb}}]^\top$. For further details, we refer the reader to \cite{Oldewurtel2011}.

The model dynamics are given below. The bilinearity is present in both input-state, and input-disturbance.
\begin{equation}\label{E:bilinear1}
x(k+1) = Ax(k)+(B_u +B_{xu}[x(k)] + B_{du}[d(k)]) u(k)+B_d d(k),
\end{equation}
where $x_{k} \in \mathbb{R}^{12}, u_{k} \in \mathbb{R}^{4}, d_{k} \in \mathbb{R}^{8} \ \forall k = 0,\dots,T$, and the matrices $B_{xu}$ and $B_{du}$ are defined as
\begin{equation}
	\begin{aligned}
		&B_{xu}[x(k)] = [ B_{xu,1}x(k),B_{xu,2}x(k), B_{xu,3}x(k), B_{xu,4}x(k) ] \in \mathbb{R}^{12\times4}, \\
		&B_{du}[d(k)] = [ B_{du,1}d(k),B_{du,2}d(k), B_{xu,3}x(k), B_{du,4}d(k) ] \in \mathbb{R}^{12\times4}, \\
	\end{aligned}
\end{equation}
with $B_{xu,i} \in \R^{12\times12}, B_{du,i} \in \R^{12\times8} \ \forall i=1,2,3,4$.
For this study, we assume that the disturbances are precisely known to MPC as well as DPC controller.

### 2018-01-20 paper/final/cs-mpc.tex
small changes

The objective of the controller is to minimize the energy usage $c^Tu$ while maintaining a desired level of thermal comfort $x_{ref}$ (or $\tT_\mathrm{ref}$).
Therefore, at time step $k$, we solve a continuously linearized MPC problem to determine the optimal sequence of inputs $[u_{\mathrm{k|k}},\dots,u_{\mathrm{k+N-1|k}}]$:

\begin{subequations}
	\begin{align}
		& \underset{u_{k+j},\epsilon_j}{\text{minimize}} & & \sum_{j=1}^{N} ({x}_{k+j} - x_{ref})\top Q ({x}_{k+j} - x_{ref}) + c^Tu_{k+j-1} +  \lambda\epsilon_j \\
		& \text{subject to }                             & & x_{k+j} = Ax_{k+j-1} + B u_{k+j-1} + B_d d_{k+j-1} \label{SE:mpc1}                                   \\
		&                                                & & B       = B_u + B_{xu}[x_{k}] + B_{du}[d_{k+j-1}] \label{SE:mpc2}                                    \\
		&                                                & & \underline{u} \leq u_{k+j-1} \leq \bar{u} 									   	                      \\
		&                                                & & \underline{x} - \epsilon_j \leq x_{k+j} \leq \bar{x} + \epsilon_j                                    \\
		&                                                & & \epsilon_j \geq 0, \ j = 1,\dots,N,            									                 
	\end{align}\label{E:mpc}
\end{subequations}

>>>>

The objective of the controller is to minimize the energy usage $c^\top u$ while maintaining a desired level of thermal comfort $x_{ref}$ (or $\tT_\mathrm{ref}$).
Therefore, at time step $k$, we solve a continuously linearized MPC problem to determine the optimal sequence of inputs $u^*_{k},\dots,u^*_{k+N-1}$:

\begin{subequations}
	\begin{align}
		& \underset{u_{k+j-1},\epsilon_j}{\text{minimize}} & & \sum_{j=1}^{N+1} ({x}_{k+j} - x_{ref})^\top Q ({x}_{k+j} - x_{ref}) + c^\top u_{k+j-1} +  \lambda\epsilon_j                                                                                                                          \\
		& \text{subject to }                               & & x_{k+j} = Ax_{k+j-1} + B u_{k+j-1} + B_d d_{k+j-1}\label{SE:mpc1}                   \\
		&                                                  & & B       = B_u + B_{xu}[x_{k}] + B_{du}[d_{k+j-1}] \label{SE:mpc2}                   \\
		&                                                  & & \underline{u} \leq u_{k+j-1} \leq \bar{u}                                           \\
		&                                                  & & \underline{x} - \epsilon_j \leq x_{k+j} \leq \bar{x} + \epsilon_j                   \\
		&                                                  & & x_k = x(k)                                                                          \\
		&                                                  & & \epsilon_j \geq 0, \ j = 1,\dots,N+1,            									                 
	\end{align}\label{E:mpc}
\end{subequations}

### 2018-01-20 paper/final/dpc.tex
small changes

\textcolor[rgb]{0,0,1}{Now, in order to have a model that can be used for prediction over an arbitrary long horizon in an MPC problem, our goal is to predict, at time $k$, the output $y$ for next $N$ time steps, i.e. $y(k),\ldots,y(k+N)$, where $N$ is the control horizon, given the state measurements, i.e. $x(k)$, and the disturbance forecast, i.e. $\tilde d(k),\ldots,\tilde d(k+N)$, at time $k$.
For the sake of simplicity and without any loss of generality, in the following, we consider only a single output, i.e. $y(k)\in\mathbb{R}$ (or $n_y = 1$).
However, as we will see in Sections \ref{S:casestudy} and \ref{S:realCaseStudy}, multiple trees (or forests) can be built for multiple outputs.
Applying the dataset splitting, we build $N$ regression trees using CART procedure \cite{Breiman1984} (see also Appendix for more details), such that the output $y(k+j)$ of the $j^{th}$ tree depends upon the previous $\delta_d$ disturbances, and on the state at instant $k$ and its $\delta_x$ regressive terms:
\begin{equation}\label{E:model_tree}
\T_j = \mathit{f}_{\mathrm{tree}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=1,\ldots,N,
\end{equation}
with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{tree}}$ representing the tree structure.
Then, the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of the trees $\T_j,\ j=1,\ldots,N$, are defined as
\begin{equation}\label{E:model_leaf}
y(k+j) =  \beta_j [1,u(k),\ldots,u(k+j) ]^T,\ j=1,\ldots,N.
\end{equation}}
Note that the coefficients $\beta_j$ are different for each leaf.
Equation~\eqref{E:model_leaf} implies that the prediction of output $y(k+j)$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j$.
Thus, we have managed to linearize the original model dynamics via black-box modeling.
\textcolor[rgb]{0,0,1}{One of the advantages of this approach is that this two-step training is done off-line, so the time required to create the models does not affect the control execution in run-time.}

\textcolor[rgb]{0,0,1}{The problem now is: \emph{how to use this modeling framework to set up an MPC problem?} In run-time, given the disturbance forecast and the state measurements at time $k$, i.e. $\tilde d(k+j-\delta_d),\ldots,\tilde d(k+j),x(k-\delta_x),\ldots,x(k)$ given $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf} for each step $j=1,\ldots,N$ of the horizon (see first part of the pseudo code given in Algorithm \ref{A:dpcrt}), as depicted in Figure \ref{F:dpc-sepvars} (left). These models are used to setup the MPC Problem \ref{P:dpcrt}.
In particular, }when a new control action has to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC. Thus, the MPC optimization problem, considering a quadratic cost ($Q \succeq 0, R \succeq 0$) function and the general case of multiple outputs, i.e. $y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, can be formulated as:
%\begin{problem}\label{P:dpcrt}
%	\begin{equation}
%		\begin{aligned}
%		\text{min } & \sum_{j=1}^{N} ({\Y}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\X^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\X^c}_{\mathrm{k+j-1|k}} + \lambda\epsilon_j\\
%		\text{s.~t. } & \ \ \ \ \ \Y_{\mathrm{k+j|k}} =  \beta^T [1,\X^c_{\mathrm{k|k}},\dots,\X^c_{\mathrm{k+j-1|k}} ]^T \\
%		& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\X}^c \leq \X^c_{\mathrm{k+j-1|k}} \leq \bar{\X}^c\\ 
%		& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\Y}-\epsilon_j \leq \Y_{\mathrm{k+j|k}} \leq \bar{\Y} + \epsilon_j\\\
%		& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
%		\end{aligned}
%		\label{E:dpcrt}
%	\end{equation}
%\end{problem}
\begin{problem}\label{P:dpcrt}
	\begin{equation}
	\begin{aligned}
	& \underset{u_{k+j}}{\text{minimize}} & & \sum_{j=1}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
	& \text{subject to }                  & & y_{k+j}     =   \beta_j [1,u_{k},\ldots,u_{k+j} ]^\top                             \\
	&                                     & & u_{k+j}    \in  \mathcal{U}                                                        \\
	&                                     & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										 \\
	&                                     & & \epsilon_j \geq  0							                                     \\
	&                                     & & j           =    1,\ldots,N.            									         \\
	\end{aligned}
	\label{E:dpcrt}
	\end{equation}
\end{problem}

@@

\textcolor[rgb]{0,0,1}{The goal with DPC-En is to replace each tree in Algorithm~\ref{A:dpcrt} by a forest.
This can be done replacing Equations \eqref{E:model_tree} and \eqref{E:model_leaf} with the Equations \eqref{E:model_forest} and \eqref{E:model_leaf_forest} provided in the following:
\begin{equation}\label{E:model_forest}
\F_j = \mathit{f}_{\mathrm{forest}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=1,\ldots,N,
\end{equation}
with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{forest}}$ representing the forest structure.
In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\bar\Y_i)$.
Then the linear models, as functions of the inputs variables $u\in\bar\X^c_i\subset\X_c$, in each leaf of each tree of the forests $\F_j,\ j=1,\ldots,N$, are defined as
\begin{equation}\label{E:model_leaf_forest}
y(k+j) =  \Theta_{ij} [1,u(k),\ldots,u(k+j) ]^T,\ j=1,\ldots,N.
\end{equation}
Here $(\bar\X^c_i,\bar\Y_i)$ correspond to the in-bag samples for the trees.}

@@

\begin{problem}\label{P:dpcrf}
	\begin{equation}
	\begin{aligned}
	& \underset{u_{k+j}}{\text{minimize}} & & \sum_{j=1}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
	& \text{subject to }                  & & y_{k+j}      =  \hat{\Theta}_j [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
	&                                     & & u_{k+j}    \in  \mathcal{U}                                                        \\
	&                                     & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										 \\
	&                                     & & \epsilon_j \geq  0							                                     \\
	&                                     & & j           =    1,\ldots,N.            									         \\
	\end{aligned}
	\label{E:dpcrf}
	\end{equation}
\end{problem}
DPC-En is graphically described in Figure~\ref{F:dpc-algo-rf}.

>>>>

\textcolor[rgb]{0,0,1}{Now, in order to have a model that can be used for prediction over an arbitrary long horizon in an MPC problem, our goal is to predict, at time $k$, the output $y$ for next $N$ time steps, i.e. $y(k),\ldots,y(k+N)$, where $N$ is the control horizon, given the state measurements, i.e. $x(k)$, and the disturbance forecast, i.e. $\tilde d(k),\ldots,\tilde d(k+N)$, at time $k$.
For the sake of simplicity and without any loss of generality, in the following, we consider only a single output, i.e. $y(k)\in\mathbb{R}$ (or $n_y = 1$).
However, as we will see in Sections \ref{S:casestudy} and \ref{S:realCaseStudy}, multiple trees (or forests) can be built for multiple outputs.
Applying the dataset splitting, we build $N$ regression trees using CART procedure \cite{Breiman1984} (see also Appendix for more details), such that the output $y(k+j)$ of the $j^{th}$ tree depends upon the previous $\delta_d$ disturbances, and on the state at instant $k$ and its $\delta_x$ regressive terms:
\begin{equation}\label{E:model_tree}
\T_j = \mathit{f}_{\mathrm{tree}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=0,\ldots,N,
\end{equation}
with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{tree}}$ representing the tree structure.
Then, the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of the trees $\T_j,\ j=0,\ldots,N$, are defined as
\begin{equation}\label{E:model_leaf}
y(k+j) =  \beta_j [1,u(k),\ldots,u(k+j) ]^T,\ j=0,\ldots,N.
\end{equation}}
Note that the coefficients $\beta_j$ are different for each leaf.
Equation~\eqref{E:model_leaf} implies that the prediction of output $y(k+j)$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j$.
Thus, we have managed to linearize the original model dynamics via black-box modeling.
\textcolor[rgb]{0,0,1}{One of the advantages of this approach is that this two-step training is done off-line, so the time required to create the models does not affect the control execution in run-time.}

\textcolor[rgb]{0,0,1}{The problem now is: \emph{how to use this modeling framework to set up an MPC problem?} In run-time, given the disturbance forecast and the state measurements at time $k$, i.e. $\tilde d(k+j-\delta_d),\ldots,\tilde d(k+j),x(k-\delta_x),\ldots,x(k)$ given $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf} for each step $j=0,\ldots,N$ of the horizon (see first part of the pseudo code given in Algorithm \ref{A:dpcrt}), as depicted in Figure \ref{F:dpc-sepvars} (left). These models are used to setup the MPC Problem \ref{P:dpcrt}.
In particular, }when a new control action has to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC. Thus, the MPC optimization problem, considering a quadratic cost ($Q \succeq 0, R \succeq 0$) function and the general case of multiple outputs, i.e. $y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, can be formulated as:

\begin{problem}\label{P:dpcrt}
	\begin{equation}
	\begin{aligned}
	& \underset{u_{k+j}}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
	& \mathrm{subject\ to }                 & & y_{k+j}     =   \beta_j [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                   \\
	&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
	&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										   \\
	&                                       & & \epsilon_j \geq  0							                                       \\
	&                                       & & j           =    0,\ldots,N.            									       \\
	\end{aligned}
	\label{E:dpcrt}
	\end{equation}
\end{problem}

@@

\textcolor[rgb]{0,0,1}{The goal with DPC-En is to replace each tree in Algorithm~\ref{A:dpcrt} by a forest.
This can be done replacing Equations \eqref{E:model_tree} and \eqref{E:model_leaf} with the Equations \eqref{E:model_forest} and \eqref{E:model_leaf_forest} provided in the following:
\begin{equation}\label{E:model_forest}
\F_j = \mathit{f}_{\mathrm{forest}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=0,\ldots,N,
\end{equation}
with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{forest}}$ representing the forest structure.
In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\bar\Y_i)$.
Then the linear models, as functions of the inputs variables $u\in\bar\X^c_i\subset\X_c$, in each leaf of each tree of the forests $\F_j,\ j=0,\ldots,N$, are defined as
\begin{equation}\label{E:model_leaf_forest}
y(k+j) =  \Theta_{ij} [1,u(k),\ldots,u(k+j) ]^T,\ j=0,\ldots,N.
\end{equation}
Here $(\bar\X^c_i,\bar\Y_i)$ correspond to the in-bag samples for the trees.}

@@

\begin{problem}\label{P:dpcrf}
	\begin{equation}
	\begin{aligned}
	& \underset{u_{k+j}}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
	& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_j [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
	&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
	&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										 \\
	&                                       & & \epsilon_j \geq  0							                                     \\
	&                                       & & j           =    0,\ldots,N.            									         \\
	\end{aligned}
	\label{E:dpcrf}
	\end{equation}
\end{problem}
DPC-En is graphically described in Figure~\ref{F:dpc-algo-rf}.
